---
title: "Supplement 1"
subtitle: "Descriptive statistics for the whole cohort"
author: "Peter Kamerman and Prinisha Pillay"
date: "`r format(Sys.Date(), '%d %B %Y')`"
---

```{r setup, include = FALSE}
# Load packages
library(magrittr)
library(tidyverse)
library(epitools)
library(boot)
library(coin)
library(skimr)

# Set plot theme
theme_set(new = theme_bw(base_size = 14))

# knitr options
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE, 
                      fig.align = 'center',
                      fig.path = './figures/basic-analysis/')
```

----

**This script generates descriptive statistics for variables collected at baseline (visit day: 0).**

The following data columns were not analysed:

- `pain` and `pain score`: related to sensory neuropathy (SN) only, therefore not relevant at baseline when everyone was free from SN.  

- `*_record`: inconsistent patient records.  

- `hba1c_percent` and `vitaminB12_pmol.l`: These data were only used to classify individuals as having diabetes or vitamin B12 deficiency when cleaning the data _(see:clean-data.R)_.  

- `ID`, `visit_number`, `visit_day`, `hivsn_present`, `visit_months`: provide sorting and grouping information only. 

----

# Import data
```{r import}
data <- readr::read_rds('data/clean_data.rds') %>% 
    # Filter for visit one
    filter(visit_number == 1) %>% 
    # Remove columns that won't be analysed
    select(-starts_with('pain'), -ends_with('_record'), 
           -visit_number, -visit_months, 
           -hba1c_percent, -vitaminB12_pmol.l) 
```

# Inspect data
```{r quick_look}
# Dimensions 
dim(data) 
# Column names
names(data)
# Head and tail
head(data)
tail(data)
# Data structure
glimpse(data)
```

# Summarise data
```{r skimr}
data %>% 
    # Convert sex to a factor
    mutate(sex = factor(sex)) %>% 
    # Remove uninformative data column
    select(-ID, -visit_day) %>% 
    skim()
```

----

# Bootstrap functions

Construct functions to calculate the 95% confidence interval of mean/median/proportion, which can be used by the `boot` package. 

```{r boot}
# Calculate mean difference between SN+ and SN-
mean_delta <- function(d, i){
    # Supply with a two-column dataframe 
    # First column is numeric variable, second column is 'sn' status
    foo <- d[i, ]
    sn <- filter(foo, sn == 'yes')
    no <- filter(foo, sn == 'no')
    sn_mean <- mean(sn[[1]], na.rm = TRUE)
    no_mean <- mean(no[[1]], na.rm = TRUE)
    delta <- no_mean - sn_mean
    delta
}

# Calculate median difference between SN+ and SN-
median_delta <- function(d, i){
    # Supply with a two-column dataframe 
    # First column is numeric variable, second column is 'sn' status
    foo <- d[i, ]
    sn <- filter(foo, sn == 'yes')
    no <- filter(foo, sn == 'no')
    sn_median <- median(sn[[1]], na.rm = TRUE)
    no_median <- median(no[[1]], na.rm = TRUE)
    delta <- no_median - sn_median
    delta
}

# Calculate odds ratio between SN+ and SN-
odds_ratio <- function(d, i){
    # Supply with a two-column dataframe 
    # First column is a dichotomous variable, second column is 'sn' status
    foo <- d[i, ]
    tab <- table(foo)
    odds_ratio <- oddsratio(tab, 
                            method = 'fisher')$measure[2, 1]
    odds_ratio
}
```

----

# Analyses

## Age
```{r age}
# Summary table (conditional on sn)
data %>%
    filter(!is.na(age_years)) %>%
    group_by(sn) %>%
    summarise(Mean = mean(age_years),
              SD = sd(age_years),
              Range = paste0(min(age_years), ' - ',
                             max(age_years)), 
              Count = n()) %>%
    knitr::kable(.,
                caption = 'Age (years) (by SN status)',
                digits = 1,
                align = 'lrrrr')

# Plots
## Univariate
data %>%
    ggplot(data = .) +
    aes(y = age_years, 
        x = 'All patients') +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'Age at recruitment',
         subtitle = '(All participants)') +
    theme_bw() +
    theme(axis.text.x = element_blank())

## Conditional on SN
data %>%
    ggplot(data = .) +
    aes(y = age_years, 
        x = sn) +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'Age at recruitment',
         subtitle = '(SN+ vs SN-)') +
    theme_bw() 

# Stats
set.seed(1234)
## Significance test
## (Fisher-Pitman Permutation Test)
oneway_test(age_years ~ factor(sn), 
            data = data,
            distribution = 'exact')

## 95% confidence interval of the mean difference
## (Bootstrap CI using bias-corrected accelerated method)
boot.ci(
    boot(data = select(data, age_years, sn), 
         statistic = mean_delta, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')
```

## Body mass
```{r mass}
# Summary table
data %>%
    filter(!is.na(mass_kg)) %>%
    summarise(Mean = mean(mass_kg),
              SD = sd(mass_kg),
              Range = paste0(min(mass_kg), ' - ',
                             max(mass_kg)),  
              Count = n()) %>%
    knitr::kable(.,
                caption = 'Body mass (kg) (whole cohort)',
                digits = 1,
                align = 'rrrr')

# Summary table (conditional on sn)
data %>%
    filter(!is.na(mass_kg)) %>%
    group_by(sn) %>%
    summarise(Mean = mean(mass_kg),
              SD = sd(mass_kg),
              Range = paste0(min(mass_kg), ' - ',
                             max(mass_kg)), 
              Count = n()) %>%
    knitr::kable(.,
                caption = 'Body mass (kg) (by SN status)',
                digits = 1,
                align = 'llrrrr')

# Plots
## Univariate
data %>%
    ggplot(data = .) +
    aes(y = mass_kg, 
        x = 'All patients') +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'Body mass at recruitment',
         subtitle = '(All participants)') +
    theme_bw() +
    theme(axis.text.x = element_blank())

## Conditional on SN
data %>%
    ggplot(data = .) +
    aes(y = mass_kg, 
        x = sn) +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'Body mass at recruitment',
         subtitle = '(SN+ vs SN-)') +
    theme_bw() 

## Conditional on SN and Sex
data %>%
    ggplot(data = .) +
    aes(y = mass_kg, 
        x = sn) +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'Body mass at recruitment, grouped by sex',
         subtitle = '(SN+ vs SN-)') +
    facet_grid(. ~ sex) +
    theme_bw() 

## Possible sex influence

# Stats
set.seed(1234)
## Significance test
## (Fisher-Pitman Permutation Test)
oneway_test(mass_kg ~ factor(sn), 
            data = data,
            distribution = 'exact')

## 95% confidence interval of the mean difference
## (Bootstrap CI using bias-corrected accelerated method)
boot.ci(
    boot(data = select(data, mass_kg, sn), 
         statistic = mean_delta, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')
```

## Height
Expect height to show sex difference, so analyse seperately for males and females.
```{r height}
# Summary table
data %>%
    filter(!is.na(height_m)) %>%
    group_by(sex) %>% # expect sex difference
    summarise(Mean = mean(height_m),
              SD = sd(height_m),
              Range = paste0(min(height_m), ' - ',
                             max(height_m)),
              Count = n()) %>%
    knitr::kable(.,
                caption = 'Height (m), grouped by sex (whole cohort)',
                digits = 2,
                align = 'lrrrr')

# Summary table (conditional on sn)
data %>%
    filter(!is.na(height_m)) %>%
    group_by(sn, sex) %>%
    summarise(Mean = mean(height_m),
              SD = sd(height_m),
              Range = paste0(min(height_m), ' - ',
                             max(height_m)),
              Count = n()) %>%
    knitr::kable(.,
                caption = 'Height (m), grouped by sex (by SN status)',
                digits = 2,
                align = 'llrrrr')

# Plots
## Univariate
data %>%
    ggplot(data = .) +
    aes(y = height_m, 
        x = sex) +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'Height at recruitment, grouped by sex',
         subtitle = '(All participants)') +
    theme_bw() 

## Conditional on SN
data %>%
    ggplot(data = .) +
    aes(y = height_m, 
        x = sn) +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'Height at recruitment, grouped by sex',
         subtitle = '(SN+ vs SN-)') +
    facet_grid(. ~ sex) +
    theme_bw() 

# Stats
set.seed(1234)
## Significance test
## (Fisher-Pitman Permutation Test)
### Male vs female (all)
oneway_test(height_m ~ factor(sex), 
            data = data,
            distribution = 'exact')
### Females (by sn status)
data %>%
    filter(sex == 'F') %>%
    oneway_test(height_m ~ factor(sn), 
                data = .,
                distribution = 'exact')
### Male (by sn status)
data %>%
    filter(sex == 'M') %>%
    oneway_test(height_m ~ factor(sn), 
                data = .,
                distribution = 'exact')

## 95% confidence interval of the mean difference
## (Bootstrap CI using bias-corrected accelerated method)
### Females 
boot.ci(
    boot(data = select(data[data$sex == 'F', ], height_m, sn), 
         statistic = mean_delta, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')
### Males 
boot.ci(
    boot(data = select(data[data$sex == 'M', ], height_m, sn), 
         statistic = mean_delta, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')
```

## Sex
```{r sex}
# Summary table
data %>%
    filter(!is.na(sex)) %>%
    group_by(sex) %>%
    summarise(Count = n(),
              Proportion = round(Count / nrow(.), 2)) %>%
    knitr::kable(.,
                caption = 'Sex frequency (whole cohort)',
                digits = 2)

# Summary table (conditional on sn)
data %>%
    filter(!is.na(sex)) %>%
    group_by(sn, sex) %>%
    summarise(Count = n(),
              Proportion = round(Count / nrow(.), 2)) %>%
    knitr::kable(.,
                caption = 'Sex frequency (by SN status)',
                digits = 2)
# Plots
## Univariate
data %>%
    ggplot(data = .) +
    aes(x = visit_number,
        fill = sex) +
    geom_bar(position = 'fill') + 
    geom_hline(yintercept = 0.5,
               linetype = 2) +
    labs(title = 'Sex ratio at recruitment',
         subtitle = '(All participants)',
         y = 'proportion') +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank())

## Conditional on SN
data %>%
    ggplot(data = .) +
    aes(x = sn,
        fill = sex) +
    geom_bar(position = 'fill') +
    geom_hline(yintercept = 0.5,
               linetype = 2) +
    labs(title = 'Sex ratio at recruitment',
         subtitle = '(SN+ vs SN-)',
         y = 'proportion') +
    theme_bw() 

# Stats
set.seed(1234)
## Significance test
## (Chi-squared test)
chisq_test(factor(sex) ~ factor(sn), 
           data = data,
           distribution = 'exact')

## 95% confidence interval of the odds ratio, caculated using 
## the conditional maximum likelihood estimation (Fisher) method
## (reference group: Female)
## (Bootstrap CI using bias-corrected accelerated method)
boot.ci(
    boot(data = select(data, sex, sn), 
         statistic = odds_ratio, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')
```

## CD4 T-cell count
```{r cd4}
# Summary table
data %>%
    filter(!is.na(CD4_cell.ul)) %>%
    summarise(Median = median(CD4_cell.ul),
              IQR = paste0(quantile(CD4_cell.ul, 0.25), ' - ',
                           quantile(CD4_cell.ul, 0.75)),
              Range = paste0(min(CD4_cell.ul), ' - ',
                             max(CD4_cell.ul)), 
              Count = n()) %>%
    knitr::kable(.,
                caption = 'CD4 T-cell count (cell/ul) (whole cohort)',
                digits = 1,
                align = 'rrrr')

# Summary table (conditional on sn)
data %>%
    filter(!is.na(CD4_cell.ul)) %>%
    group_by(sn) %>%
    summarise(Median = median(CD4_cell.ul),
              IQR = paste0(round(quantile(CD4_cell.ul, 0.25)), ' - ',
                           round(quantile(CD4_cell.ul, 0.75))),
              Range = paste0(min(CD4_cell.ul), ' - ',
                             max(CD4_cell.ul)), 
              Count = n()) %>%
    knitr::kable(.,
                caption = 'CD4 T-cell count (cell/ul) (by SN status)',
                digits = 0,
                align = 'lrrrr')

# Plots
## Univariate
data %>%
    ggplot(data = .) +
    aes(y = CD4_cell.ul, 
        x = 'All patients') +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'CD4 T-cell count at recruitment',
         subtitle = '(All participants)') +
    theme_bw() +
    theme(axis.text.x = element_blank())

## Conditional on SN
data %>%
    ggplot(data = .) +
    aes(y = CD4_cell.ul, 
        x = sn) +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'CD4 T-cell count at recruitment',
         subtitle = '(SN+ vs SN-)') +
    theme_bw() 

# Stats
set.seed(1234)
## Significance test
## (Fisher-Pitman Permutation Test)
oneway_test(CD4_cell.ul ~ factor(sn), 
            data = data,
            distribution = 'exact')

## 95% confidence interval of the median difference
## (Bootstrap CI using bias-corrected accelerated method)
boot.ci(
    boot(data = select(data, CD4_cell.ul, sn), 
         statistic = median_delta, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')
```

## Viral load
```{r viral_load}
# Summary table
data %>%
    filter(!is.na(viral_load_copies.ml)) %>%
    summarise(Median = median(viral_load_copies.ml),
              IQR = paste0(round(quantile(viral_load_copies.ml, 0.25), 2), 
                           ' - ',
                           round(quantile(viral_load_copies.ml, 0.75), 2)),
              Range = paste0(round(min(viral_load_copies.ml), 2), ' - ',
                             round(max(viral_load_copies.ml), 2)), 
              Count = n()) %>%
    knitr::kable(.,
                caption = 'Viral load [log10(copies/ml)] (whole cohort)',
                digits = 2,
                align = 'rrrr')

# Summary table (conditional on sn)
data %>%
    filter(!is.na(viral_load_copies.ml)) %>%
    group_by(sn) %>%
    summarise(Median = median(viral_load_copies.ml),
              IQR = paste0(round(quantile(viral_load_copies.ml, 0.25), 2), 
                           ' - ',
                           round(quantile(viral_load_copies.ml, 0.75), 2)),
              Range = paste0(round(min(viral_load_copies.ml), 2), ' - ',
                             round(max(viral_load_copies.ml), 2)), 
              Count = n()) %>%
    knitr::kable(.,
                caption = 'Viral load [log10(copies/ml)] (by SN status)',
                digits = 1,
                align = 'lrrrr')

# Plots
## Univariate
data %>%
    filter(!is.na(viral_load_copies.ml)) %>%
    ggplot(data = .) +
    aes(y = viral_load_copies.ml, 
        x = 'All patients') +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'CD4 T-cell count at recruitment',
         subtitle = '(All participants)',
         y = 'log10(viral_load_copies.ml)') +
    theme_bw() +
    theme(axis.text.x = element_blank())

## Conditional on SN
data %>%
    filter(!is.na(viral_load_copies.ml)) %>%
    ggplot(data = .) +
    aes(y = log10(viral_load_copies.ml), 
        x = sn) +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'CD4 T-cell count at recruitment',
         subtitle = '(SN+ vs SN-)',
         y = 'log10(viral_load_copies.ml)') +
    theme_bw() 

# Stats
set.seed(1234)
## Significance test
## (Fisher-Pitman Permutation Test)
data %>%
    filter(!is.na(viral_load_copies.ml)) %>%
    oneway_test(viral_load_copies.ml ~ factor(sn), 
                data = .)

## 95% confidence interval of the median difference
## (Bootstrap CI using bias-corrected accelerated method)
data_log.vl <- data %>%
    filter(!is.na(viral_load_copies.ml)) %>%
    select(viral_load_copies.ml, sn)

boot.ci(
    boot(data = data_log.vl, 
         statistic = median_delta, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')

# Clean-up
rm(data_log.vl)
```

## Alcohol
### Consumes alcohol, yes or no?
```{r alcohol}
# Summary table
data %>%
    filter(!is.na(consumes_alcohol)) %>%
    group_by(consumes_alcohol) %>%
    summarise(Count = n(),
              Proportion = round(Count / nrow(.), 2)) %>%
    knitr::kable(.,
                caption = 'Consumes alcohol (whole cohort)',
                digits = 2)

# Summary table (conditional on sn)
data %>%
    filter(!is.na(consumes_alcohol)) %>%
    group_by(sn, consumes_alcohol) %>%
    summarise(Count = n(),
              Proportion = round(Count / nrow(.), 2)) %>%
    knitr::kable(.,
                caption = 'Consumes alcohol (by SN status)',
                digits = 2)
# Plots
## Univariate
data %>%
    ggplot(data = .) +
    aes(x = visit_number,
        fill = consumes_alcohol) +
    geom_bar(position = 'fill') + 
    geom_hline(yintercept = 0.5,
               linetype = 2) +
    labs(title = 'Consumes alcohol at recruitment',
         subtitle = '(All participants)',
         y = 'proportion') +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank())

## Conditional on SN
data %>%
    ggplot(data = .) +
    aes(x = sn,
        fill = consumes_alcohol) +
    geom_bar(position = 'fill') +
    geom_hline(yintercept = 0.5,
               linetype = 2) +
    labs(title = 'Consumes alcohol at recruitment',
         subtitle = '(SN+ vs SN-)',
         y = 'proportion') +
    theme_bw() 

# Stats
set.seed(1234)
## Significance test
## (Chi-squared test)
chisq_test(factor(consumes_alcohol) ~ factor(sn), 
           data = data,
           distribution = 'exact')

## 95% confidence interval of the odds ratio, caculated using 
## the conditional maximum likelihood estimation (Fisher) method
## (reference group: No alcohol)
## (Bootstrap CI using bias-corrected accelerated method)
boot.ci(
    boot(data = select(data, consumes_alcohol, sn), 
         statistic = odds_ratio, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')
```

### Units of alcohol consumed per week
_(If a person consumes alcohol)_
```{r units_alcohol}
data_eoh <- data %>%
    filter(consumes_alcohol == 'yes') %>%
    select(alcohol_units.week, sn)

# Summary table
data_eoh %>%
    summarise(Median = median(alcohol_units.week),
              IQR = paste0(round(quantile(alcohol_units.week, 0.25), 2), 
                           ' - ',
                           round(quantile(alcohol_units.week, 0.75), 2)),
              Range = paste0(round(min(alcohol_units.week), 2), ' - ',
                             round(max(alcohol_units.week), 2)), 
              Count = n()) %>%
    knitr::kable(.,
                caption = 'Alcohol consumed (units/week) (whole cohort)',
                digits = 2,
                align = 'rrrr')

# Summary table (conditional on sn)
data_eoh %>%
    group_by(sn) %>%
    summarise(Median = median(alcohol_units.week),
              IQR = paste0(round(quantile(alcohol_units.week, 0.25), 2), 
                           ' - ',
                           round(quantile(alcohol_units.week, 0.75), 2)),
              Range = paste0(round(min(alcohol_units.week), 2), ' - ',
                             round(max(alcohol_units.week), 2)), 
              Count = n()) %>%
    knitr::kable(.,
                caption = 'Alcohol consumed (units/week) (by SN status)',
                digits = 1,
                align = 'lrrrr')

# Plots
## Univariate
data_eoh %>%
    ggplot(data = .) +
    aes(y = alcohol_units.week, 
        x = 'All patients') +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'Alcohol consumed (units/week) at recruitment',
         subtitle = '(All participants)') +
    theme_bw() +
    theme(axis.text.x = element_blank())

## Conditional on SN
data_eoh %>%
    ggplot(data = .) +
    aes(y = alcohol_units.week, 
        x = sn) +
    geom_boxplot() +
    geom_jitter(height = 0) +
    labs(title = 'Alcohol consumed (units/week) at recruitment',
         subtitle = '(SN+ vs SN-)') +
    theme_bw() 

# Stats
set.seed(1234)
## Significance test
## (Fisher-Pitman Permutation Test)
data_eoh %>%
    oneway_test(alcohol_units.week ~ factor(sn), 
                data = .,
                distribution = 'exact')

## 95% confidence interval of the median difference
## (Bootstrap CI using bias-corrected accelerated method)
boot.ci(
    boot(data = data_eoh, 
         statistic = median_delta, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')

# Clean-up
rm(data_eoh)
```

## TB infection/treatment
### Do you currently have TB?
```{r tb_current}
# Summary table
data %>%
    filter(!is.na(TB_current)) %>%
    group_by(TB_current) %>%
    summarise(Count = n(),
              Proportion = round(Count / nrow(.), 2)) %>%
    knitr::kable(.,
                caption = 'Currently has TB infection (whole cohort)',
                digits = 2)

# Summary table (conditional on sn)
data %>%
    filter(!is.na(TB_current)) %>%
    group_by(sn, TB_current) %>%
    summarise(Count = n(),
              Proportion = round(Count / nrow(.), 2)) %>%
    knitr::kable(.,
                caption = 'Currently has TB infection (by SN status)',
                digits = 2)
# Plots
## Univariate
data %>%
    ggplot(data = .) +
    aes(x = visit_number,
        fill = TB_current) +
    geom_bar(position = 'fill') + 
    geom_hline(yintercept = 0.5,
               linetype = 2) +
    labs(title = 'Current TB infection at recruitment',
         subtitle = '(All participants)',
         y = 'proportion') +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank())

## Conditional on SN
data %>%
    ggplot(data = .) +
    aes(x = sn,
        fill = TB_current) +
    geom_bar(position = 'fill') +
    geom_hline(yintercept = 0.5,
               linetype = 2) +
    labs(title = 'Current TB infection at recruitment',
         subtitle = '(SN+ vs SN-)',
         y = 'proportion') +
    theme_bw() 

# Stats
set.seed(1234)
## Significance test
## (Chi-squared test)
chisq_test(factor(TB_current) ~ factor(sn), 
           data = data,
           distribution = 'exact')

## 95% confidence interval of the odds ratio, caculated using 
## the conditional maximum likelihood estimation (Fisher) method
## (reference group: No TB)
## (Bootstrap CI using bias-corrected accelerated method)
boot.ci(
    boot(data = select(data, TB_current, sn), 
         statistic = odds_ratio, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')
```

### Currently receiving TB treatment?

Treatment consists of rifafour and pyridoxine (prophylaxis). Therefore only need to analyse rifafour data. _Note: Treatment policy was to start some patients, irrespective of TB diagnosis, on TB treatment._

```{r tb_rifafour}
# Double-check matching between rifafour and pyridoxine columns 
unique(data$rifafour_treatment == data$pyridoxine_prophylaxis)

# Summary table
data %>%
    filter(!is.na(rifafour_treatment)) %>%
    group_by(rifafour_treatment) %>%
    summarise(Count = n(),
              Proportion = round(Count / nrow(.), 2)) %>%
    knitr::kable(.,
                caption = 'Currently receiving rifafour/pryridoxine (whole cohort)',
                digits = 2)

# Summary table (conditional on sn)
data %>%
    filter(!is.na(rifafour_treatment)) %>%
    group_by(sn, rifafour_treatment) %>%
    summarise(Count = n(),
              Proportion = round(Count / nrow(.), 2)) %>%
    knitr::kable(.,
                caption = 'Currently receiving rifafour/pryridoxine (by SN status)',
                digits = 2)
# Plots
## Univariate
data %>%
    ggplot(data = .) +
    aes(x = visit_number,
        fill = rifafour_treatment) +
    geom_bar(position = 'fill') + 
    geom_hline(yintercept = 0.5,
               linetype = 2) +
    labs(title = 'Receiving rifafour/pryridoxine at recruitment',
         subtitle = '(All participants)',
         y = 'proportion') +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank())

## Conditional on SN
data %>%
    ggplot(data = .) +
    aes(x = sn,
        fill = rifafour_treatment) +
    geom_bar(position = 'fill') +
    geom_hline(yintercept = 0.5,
               linetype = 2) +
    labs(title = 'Receiving rifafour/pryridoxine at recruitment',
         subtitle = '(SN+ vs SN-)',
         y = 'proportion') +
    theme_bw() 

# Stats
set.seed(1234)
## Significance test
## (Chi-squared test)
chisq_test(factor(rifafour_treatment) ~ factor(sn), 
           data = data,
           distribution = 'exact')

## 95% confidence interval of the odds ratio, caculated using 
## the conditional maximum likelihood estimation (Fisher) method
## (reference group: No rifafour)
## (Bootstrap CI using bias-corrected accelerated method)
boot.ci(
    boot(data = select(data, rifafour_treatment, sn), 
         statistic = odds_ratio, 
         R = 1000, 
         stype = 'i'),
    type = 'bca')
```

## Diabetes

Classed as diabetic based on `data$hba1c_percent` > 7%. No one was diabetic.

```{r diabetes}
# Summary table
data %>%
    filter(!is.na(diabetic_hba1c)) %>%
    group_by(diabetic_hba1c) %>%
    summarise(Count = n(),
              Proportion = round(Count / nrow(.), 2)) %>%
    knitr::kable(.,
                caption = 'Diabetic (HBA1c > 7%) (whole cohort)',
                digits = 2)
```

## Vitamin B12 deficiency

Classed as B12 deficient based on `data$vitaminB12_pmol.l` < 141 pmol/l. Only one person had a deficiency.

```{r vitaminB12}
# Summary table
data %>%
    filter(!is.na(vitaminB12_deficiency)) %>%
    group_by(vitaminB12_deficiency) %>%
    summarise(Count = n(),
              Proportion = round(Count / nrow(.), 2)) %>%
    knitr::kable(.,
                caption = 'B12 deficiency (B12 < 141 pmol/l) (whole cohort)',
                digits = 2)

# Did the persson with a deficiency develop sn?
data[data$vitaminB12_deficiency == 'yes' & !is.na(data$vitaminB12_deficiency),
      c('vitaminB12_deficiency', 'sn')] %>%
    knitr::kable(.)
```

# Session information
```{r session_info, echo = FALSE}
# clean-up
rm(list = ls())

# Session information
sessionInfo()
```

